# Data filtering and wrangling

In this section, we will learn to use a few functions to filter, select and re-organize a data object.

## filter()

**filter()** will filter (keep or remove) **rows**.

We can start from **geneexp** object, and filter only down-regulated genes using the **filter()** function of **{dplyr}**.

If you need to import the data again, run the following command:

```{r, eval=FALSE}
geneexp <- read_csv("DataViz_source_files-main/files/expression_20genes.csv")
```

```{r}
filter(geneexp, DE=="Down")
```

**==** is a **logical operator** that represents equality. It means that filter will return **rows in geneexp that are exactly equal to "Down"**.

Logical operators:

| Operator  | Description   |
|:--------: |:-------------:|
|     <     | less than    |
|     <=    | less than or equal to   |
|     >     | greater than  |
|     >=    | greater than or equal to|
|     ==    | exactly equal to   |
|     !=    | not equal to |
|    !x     | not x |
|    x \| y  | x OR y |
|    x & y  | x AND y |

We can have several conditions.

For example, we may want to extract only rows that have either **Up** OR **Down** in **DE**:

```{r}
filter(geneexp, DE=="Down" | DE=="Up")
```

Here, we introduce another operator, **|**, that means **OR**: rows are kept if values are either **Down** OR **Up** in **DE** column.

A good practice is to assign the filtered output to a new object, for example:

```{r}
geneexp_filt <- filter(geneexp, DE=="Down" | DE=="Up")
```

## select()

**select()** will select (keep or remove) **columns** from a data frame.

Select only columns **Gene** and **DE** from **geneexp**:

```{r}
select(geneexp, Gene, DE)
```

Remove column **DE**:

```{r}
select(geneexp, -DE)
```

Again, a good practice is to save that subset into a new object:

```{r}
geneexp_selected <- select(geneexp, Gene, DE)
```

## rename()

You can rename a column using the **rename()** function.

For example, we can rename column **Gene** to **Symbol**, the following way:

```{r}
rename(geneexp, Symbol=Gene)
```

The first argument is the data, then comes the column(s) to rename in the format: *new_name=old_name*.


## Exercise 4

In this short exercise, you will practice using the functions for data manipulation we just learned.

We will start from **gtf**.

If **gtf** it is no longer in your environment, load it with:

```{r, eval=F}
gtf <- read_csv("DataViz_source_files-main/files/gencode.v44.annotation.csv")
```

1. Rename column **chr** as **Chromosome**. Assign to a new object.

<details>
<summary>
correction
</summary>

```{r}
gtf2 <- rename(gtf, chromosome=chr)
```

</details>

<br>
2. Keep rows with **"+"** strand on chromosome **"chr4"**. Assign this filtered data to a new object.

<details>
<summary>
correction
</summary>

```{r}
gtf3 <- filter(gtf2, strand=="+" & chromosome=="chr4")
```

</details>

<br>

3. Remove columns **strand** and **gene_type**.

<details>
<summary>
correction
</summary>

```{r}
gtf4 <- select(gtf3, -strand, -gene_type)
```

</details>

<br>

