# Exercises to do at home

The objective of the following exercises is to help you practice what was learned so far during the workshop. You can practice at home and reach out if you have any issue.
<br>

Exercise 5: barplot
<br>
Exercise 6: scatter plot
<br>
Exercise 7: boxplot
<br>

## Exercise 5: barplot

1. Import **DataViz_source_files-main/files/population_age_group_per_sex_long_format.csv** into an object called **pop_age**.

<details>
<summary>
correction
</summary>

```{r}
pop_age <- read_csv("DataViz_source_files-main/files/population_age_group_per_sex_long_format.csv")
```

</details>

<br>

2. Create a barplot that shows the population count per **Year**:

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=pop_age, mapping=aes(x=Year, y=Population)) + 
  geom_bar(stat="identity")
```

</details>

<br>

3. Split boxes per Sex (**in the same plot**: do not split into different plots):


<details>
<summary>
correction
</summary>

```{r}
ggplot(data=pop_age, mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity")
```

</details>

<br>

4. Keep only data that corresponds to "Spain". You can do this in 2 steps: filter first, then plot.

  
<details>
<summary>
correction
</summary>

```{r}
pop_age_spain <- filter(pop_age, Country=="Spain")
ggplot(data=pop_age_spain, mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity")
```


</details>

<br>

5. Display the bars next to each other:

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=pop_age_spain, mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity", position="dodge")
```

</details>

<br>


6. Change default colors using **scale_fill_manual()** layer.

The structure is not easy to remember: take a peek at the answer!

  
<details>
<summary>
correction
</summary>

```{r}
ggplot(data=pop_age_spain, mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=c(Female="darkgreen", Male="red"))
```

</details>

<br>

7. Add a title, change the theme / background:

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=pop_age_spain, mapping=aes(x=Year, y=Population, fill=Sex)) +
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=c("darkgreen", "red")) + 
  ggtitle("Spanish population from 1950 to 2021 per sex") + 
  theme_minimal()
```

</details>

<br>
8. Create one plot per **Age group** using **facet_wrap()**

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=pop_age_spain, mapping=aes(x=Year, y=Population, fill=Sex)) +
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=c("darkgreen", "red")) + 
  ggtitle("Spanish population from 1950 to 2021 per sex") + 
  theme_minimal() +
  facet_wrap(~Age_group)
```

</details>
<br>

9. Save to a PDF file: Export -> Save as PDF. Note: when saving to PDF format, you cannot manualy drag the plot to get the size and proportion you want, but you can **Preview** before saving it.

If you prefer, you can save the file using **ggsave()** (or you can do both!).

<details>
<summary>
correction
</summary>

```{r}
# first, save to an object
plot_age <- ggplot(data=pop_age_spain, mapping=aes(x=Year, y=Population, fill=Sex)) +
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=c("darkgreen", "red")) + 
  ggtitle("Spanish population from 1950 to 2021 per sex") + 
  theme_minimal() +
  facet_wrap(~Age_group)

# save file
ggsave(filename="population_age_sex_Spain.pdf", plot=plot_age, device = "pdf", units = "in", width = 10, height = 8)
```

</details>

<br>

10. Save your filtered data as a **.csv** file: have a look at the **write_csv()** function from {readr}:

<details>
<summary>
correction
</summary>

```{r}
write_csv(pop_age_spain, "subset_spain.csv")
```

</details>

## Exercise 6: scatter plot

1. Import **DataViz_source_files-main/files/GSE150029_rnaseq_log2.csv** into an object called **rnaseq**.

<details>
<summary>
correction
</summary>

```{r}
rnaseq <- read_csv("DataViz_source_files-main/files/GSE150029_rnaseq_log2.csv")
```

</details>
<br>

2. Create a scatter plot of sample CTRL (x-axis) versus sample EZH (y-axis).

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=rnaseq, mapping=aes(x=CTRL, y=EZH)) +
  geom_point()
```

</details>

<br>
3. Map/color the points to the **gene_biotype**

<details>
<summary>
correction
</summary>

```{r, fig.width=16}
ggplot(data=rnaseq, mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point()
```

</details>

<br>
3. Not very readable! Filter and plot only data corresponding to **either lincRNA OR miRNA**.

<details>
<summary>
correction
</summary>

```{r}
rnaseq_linc_mirna_filter1 <- filter(rnaseq, (gene_biotype=="lincRNA" | gene_biotype=="miRNA"))

ggplot(data=rnaseq_linc_mirna_filter1, mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point()
```

</details>

<br>
4. Keep only those lincRNAs and miRNAs expressed in CTRL **at least 1.5 times more than in EZH**.

<details>
<summary>
correction
</summary>

```{r}
rnaseq_linc_mirna_filter2 <- filter(rnaseq, (gene_biotype=="lincRNA" | gene_biotype=="miRNA") & CTRL > 1.5*EZH)

ggplot(data=rnaseq_linc_mirna_filter2, mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point()
```

</details>

5. Add a title to the plot, and make it **bold** (see **theme()** section of the course)

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=rnaseq_linc_mirna_filter2, mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point() +
  ggtitle("lincRNA and miRNA") +
  theme(plot.title = element_text(face = "bold"))
```

</details>
<br>

<details>
<summary>
NOTE: If you want to label only one (or few) point(s), click here.
</summary>

First, filter the data frame:
```{r}
SNHG8 <- filter(rnaseq, gene_name=="SNHG8")
```

Then, add it to geom_text:

```{r}
ggplot(data=rnaseq_linc_mirna_filter2, mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point() +
  ggtitle("lincRNA and miRNA") +
  theme(plot.title = element_text(face = "bold")) +
  geom_text(data=SNHG8, label="SNHG8", show.legend = FALSE)
```

</details>
<br>


## Exercise 7: boxplot

1. Read in **DataViz_source_files-main/files/GSE150029_rnaseq_log2_long.csv** into a new object called **rnaseq2**.

<details>
<summary>
correction
</summary>

```{r}
rnaseq2 <- read_csv("DataViz_source_files-main/files/GSE150029_rnaseq_log2_long.csv")
```

</details>

<br>

2. Create a boxplot with expression/count distribution per sample.

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=rnaseq2, mapping=aes(x=sample, y=log2_counts)) + 
  geom_boxplot()
```

</details>

<br>

3. Split the boxes per **gene_biotype**.

<details>
<summary>
correction
</summary>

```{r, fig.width=18, fig.height=9}
ggplot(data=rnaseq2, mapping=aes(x=sample, y=log2_counts, fill=gene_biotype)) + 
  geom_boxplot()
```

</details>
<br>

4. Keep only *protein_coding* and *lincRNA* biotypes (you can save the filtered data into a new object) and re-do the same plot as in 3.

<details>
<summary>
correction
</summary>

```{r}
rnaseq2_filtered <- filter(rnaseq2, gene_biotype=="protein_coding" | gene_biotype=="lincRNA")

ggplot(data=rnaseq2_filtered, mapping=aes(x=sample, y=log2_counts, fill=gene_biotype)) + 
  geom_boxplot()
```

</details>

<br>

5. Add a **geom_violin()** layer. Set **alpha=0.3** in geom_violin. What is the alpha parameter?

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=rnaseq2_filtered, mapping=aes(x=sample, y=log2_counts, fill=gene_biotype)) + 
  geom_boxplot() +
  geom_violin(alpha=0.3)

# if boxplot and violin plots are misaligned, you can play with the position parameter in geom_violin, such as:
# geom_violin(position=position_dodge(0.7))
```

</details>

<br>

6. Look at the help page of **geom_boxplot()** and change the following parameters:

* Set outlier color to red
* Set outlier shape as triangles

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=rnaseq2_filtered, mapping=aes(x=sample, y=log2_counts, fill=gene_biotype)) + 
  geom_boxplot(outlier.colour = "red", outlier.shape = "triangle")
```

</details>

