# Interactive plots with {plotly}

**{plotly}** is a package that makes interactive graphs.

There are many applications to plotly, but a simple one is to convert a plot created by ggplot into an interactive version.

First we can install **plotly** (bottom-right panel -> Packages -> Install).

And then we load the package: **tick** the package in the list.

<br>

Let's read again one of our first data file:

```{r, echo=T, eval=F}
geneexp <- read_csv("DataViz_source_files-main/files/expression_20genes.csv")
```

We can first save the "ggplot" as an object **myplot**:

```{r}
myplot <- ggplot(data=geneexp, mapping=aes(x=sample1, y=sample2, label=Gene, color=DE)) + 
  geom_point()
```

And then input that object into the **ggplotly()** function. 

```{r, echo=T, eval=F}
ggplotly(p=myplot)
```

The interactive plot should appear in the **Viewer** tab (bottom-right panel). Not that you can zoom into the plot!

Customize what appears in the tooltip box. For example, here, we only show the gene label information:

```{r, echo=T, eval=F}
ggplotly(p=myplot, tooltip="label")
```

It also works for different types of plots: for example, barplots.

```{r}
mybarplot <- ggplot(data=geneexp, mapping=aes(x=DE)) + 
  geom_bar()
```

```{r, echo=T, eval=F}
ggplotly(p=mybarplot)
```

You can add ggplot2 layers / components:

```{r}
mybarplot <- ggplot(data=geneexp, mapping=aes(x=DE)) + 
  geom_bar() +
  ggtitle(label="barplot in ggplotly") +
  theme_bw()
```

```{r, echo=T, eval=F}
ggplotly(p=mybarplot)
```
